Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-06-06 paper/paper.tex
revise section 6 and 7

\begin{lstlisting}
def inject(e: Expr): State =
  State(e, Map(),
        Store[BAddr, Storable](Map()),
        Store[KAddr, Cont](Map(Halt -> Set())),
        Halt, List())
\end{lstlisting}

@@

An instance of @VS@ represents the computational result of one path in
the nondeterministic evaluation.
The reason that we include a store is that there might be a case in which two paths
have the same values but different accumulated side effects (e.g., memory allocations).
Since the whole computation is nondeterministic, the type of the final result
values is a collection @Set[VS]@. For convenience, we define it as a type @Ans@:

@@

An additional argument @continue@ which has type @Cont@ is also introduced.
Recall that in the disentangled AAM, the second case in @drive_step@
makes a call to @continue@. Here, the reemerging @continue@ is a function
with type @Ans => Ans@. The function @aeval@ does not return: instead, it calls
@continue@ with the result values.

@@

\begin{lstlisting}
def nd[T](ts: Iterable[T], acc: Ans, k: (T, Ans, Ans => Ans) => Ans, m: Ans => Ans): Ans = {
  if (ts.isEmpty) m(acc)
  else k(ts.head, acc, (ans: Ans) => nd(ts.tail, ans, k, m))
}
\end{lstlisting}

Given a collection of elements of type @T@, an initial accumlated value of type 
@Ans@, a function @k@ that works on element's type @T@ and accumulated value's 
type @Ans@ (note that @k@ also takes a continuation of type @Ans=>Ans@), 
and a function @m@ that works on and returns a value of type @Ans@, 
@nd@ chooses an element in the collection and applies @k@ to it, accumulates 
the value and finally applies @m@ on the accumulated value when there is no element 
in the collection.

If we look at @continue@ and @ndcontinue@ in the disentangled AAM, we will find
that @nd@ is an abstraction of those two functions. In function @continue@,
there are two out calls to @ndcontinue@ and @drive_step@, this is the reason why
we have functions @k@ and @m@ in @nd@.

Readers familiar with functional programming may think of @fold@ at this point; indeed,
it is an implementation of @fold@ in continuation style. But here we have two
continuations, one for the next element in the collection, and one for the final result.
This is sometimes referred to as an extended continuation-passing style (ECPS)
that function @k@ is a normal continuation and @m@ is a meta-continuation 
\cite{Danvy:1990:AC:91556.91622}.

\textbf{Evaluation without Caching.}
Since our first version of @aeval@ implements the nondeterministic
evaluation, given the defined @nd@, the evaluation function @aeval@ can be easily
defined. In the case that @e@ is a @let@ expression,
we invoke @nd@ twice and perform a nesting two-step, depth-first evaluation
over possible values of @App(f, ae)@.
The values of branches from a fork point will be accumulated, and then
returned to the upper fork point.

\begin{lstlisting}
case Let(x, App(f, ae), e) =>
  val closures = atomicEval(f, env, store)
  nd[Storable](closures, Set[VS](), { case (clos, clsacc, clsnd) =>
    val Clos(Lam(v, body), c_env) = clos
    val baddr = allocBind(v, new_time)
    val new_cenv = c_env + (v -> baddr)
    val new_store = store.update(baddr, atomicEval(ae, env, store))
    aeval(body, new_cenv, new_store, new_time, (bdvss: Set[VS]) => {
      nd[VS](bdvss, Set[VS](), { case (vs, bdacc, bdnd) =>
        val VS(vals, time, store) = vs
        val baddr = allocBind(x, time)
        val new_env = env + (x -> baddr)
        val new_store = store.update(baddr, vals)
        /* evaluates and accumluates the values of one path, then call bdnd */
        aeval(e, new_env, new_store, time, { case evss => bdnd(bdacc++evss) }) 
      }, { case evss => clsnd(evss++clsacc) }) /* accumulate the values of (multiple) paths, then call clsnd */
    })
  }, continue)
\end{lstlisting}

The outer call to @nd@ evaluates over the set of possible closures of
@f@. For each such closure, we go into its body with a new environment and a new store.
Remember that the function @aeval@ returns a set of @VS@ objects to its continuation,
so the continuation argument @bdvss@ is a set of @VS@ objects which represent values
from multiple computation paths when evaluating this single @body@ expression.

We next consider the inner call to @nd@, which evaluates over the set of body values @bdvss@.
For each @VS@, the timestamp and store in @VS@ will be instantiated to
@aeval@.
The inner application of @aeval@ returns a set of values of one computation path to its
continuation, where the continuation will accumulate the result and transfer
evaluation to the next body value.

@@

If the expression is atomic, then we simply construct a new instance of
@VS@ and return it to the @continue@ function.

At the end, the initial invocation to @aeval@ is passed with an identity
function as a halting continuation:
\begin{lstlisting}
def analyze(e: Expr) =
  aeval(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List(), (ans => ans))
\end{lstlisting}

@@

Based on the refunctionalized AAM from the last section, we use the same nondeterministic
operator @nd@, but further extend it to a fixed-point cached evaluation in this section.
The fixed-point caching algorithm performs a sound over-approximation of all possible
concrete reachable paths and values, and also prevents non-termination when analyzing 
programs that may diverge.

@@

\textbf{Cache.}
Recall that the latent assumption of the worklist algorithm in classical AAM
is that if we have seen a state @s@, it means we also have seen the successors
of state @s@.
The caching algorithm we adopt here replays the same assumption but in a big-step
manner: if we have seen the configuration @c@, then it means we also have seen
the values that are evaluated by the configuration @c@.
Here we will apply the fixed-point caching algorithm as described from
\citeauthor{darais2017abstracting}'s ADI paper \cite{darais2017abstracting}.

@@

\textbf{Cache-Passing Style.}
We can now transform our interpreter into cache-passing style.
We begin by changing the return type @Ans@ to a case class which contains a @VS@ set and
a cache.

\begin{lstlisting}
case class Ans(vss: Set[VS], cache: Cache)
\end{lstlisting}

The skeleton of function @aeval@ largely remains the same.
Every place which has a value of @Ans@ is also replaced as direct pattern matching so
that we can use the fields in @Ans@ immediately.
The cache is used in a monotonic way: each function call to @aeval@ or @nd@ is
passed with the latest cache from the most recent continuation's result.

@@

\begin{lstlisting}
def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, continue: Cont): Ans = {
  val config = Config(e, env, store, time)
  /* lookups the out cache, uses the values if contains this config */
  if (cache.outContains(config)) cont(Ans(cache.outGet(config), cache))
  else {
    val new_cache = cache.outUpdate(config, cache.inGet(config))
    val new_time = config.tick 
    e match {
      case Let(x, App(f, ae), e) =>
        val closures = atomicEval(f, env, store)
        nd[Storable](closures, Ans(Set[VS](), new_cache), {
          case (Clos(Lam(v, body), c_env), clsans, clsnd) =>
            ... /* sets vairables, e.g., new_cenv and new_store */
            aeval(body, new_cenv, new_store, new_time, clsans.cache, { case Ans(bdvss, bdcache) =>
              nd[VS](bdvss, Ans(Set[VS](), bdcache), {
                case (VS(vals, time, vssstore), bdans, bdnd) =>
                  ... /* sets variables, e.g., new_env and new_store */
                  aeval(e, new_env, new_store, time, bdans.cache, {
                    /* updates the out cache after evaluated to the end of one path  */
                    case Ans(evss, ecache) => bdnd(bdans ++ Ans(evss, ecache.outUpdate(config, evss)))
                  })
              }, { case eans => clsnd(eans ++ clsans) })
            })
        },
        continue)

      case ae if isAtomic(ae) =>
        val vs = Set(VS(atomicEval(ae, env, store), new_time, store))
        continue(Ans(vs, new_cache.outUpdate(config, vs)))
    }
  }
}
\end{lstlisting}

@@

\begin{lstlisting}
val mtEnv = Map[String, BAddr](); val mtStore = Store[BAddr, Storable](Map()); val mtTime = List()
val mtCache = Cache(Store[Config, VS](Map()), Store[Config, VS](Map()))

def analyze(e: Expr) = {
  def iter(cache: Cache): Ans = {
    val Ans(vss, new_cache) = aeval(e, mtEnv, mtStore, mtTime, cache, {
      case Ans(vss, cache) => 
        val initConfig = Config(e, mtEnv, mtStore, mtTime) 
        Ans(vss, cache.outUpdate(initConfig, vss))
    })
    if (new_cache.out == cache.out) { Ans(vss, new_cache) }
    else { iter(Cache(new_cache.out, new_cache.out)) }
  }
  iter(mtCache)
}
\end{lstlisting}

@@

\textbf{Nondeterminism Operators, Again.}
The first operator @nd@ is adapted from the previous one, but here @nd@ has only one 
continuation @k@. Introducing @Cache@ explicitly as the part of arguments of 
continuation @k@ reveals that cache should be monotonically accumulated to 
the rest of computation. Now the only continuation @k@ applies on type @(T, Cache)@ where
the value of type @T@ comes from the collection, and @nd@ performs recursive invocation
on itself over elements in the collection @ts@. We also move the accumulation operation
(@++@) inside of @nd@.

@@

The function @choices@ takes two arguments: a set of elements with type @T@, 
and an initial accumulated value of type @Ans@.
The @choices@ iteratively returns an element from the collection along with the latest cache, 
and the return type is also annotated by \verb|@cps[Ans]|. 
This annotation denotes the final returned value of @choices@ is type @Ans@.
The @f@ introduced by the @shift@ operator is the delimited continuation from the 
call-site of @choices@.

Now we can use these two operators to redefine @aeval@.
We add an annotation \verb|@cps[Ans]| on the return type of the @aeval@ function.
Then, for every nondeterministic choice in the abstract interpreter, we can simply call 
@choices@ on the set of possible choices and write the program sequentially as concrete 
interpreters, and the subsequent statements and expressions become the delimited continuation
@f@. @choices@ also returns the latest cache to its left-hand side definition.
The cache is still accumulated to the next call to @choices@ or @aeval@, 
given the fact that the abstract interpreter should always use the latest cache.

\begin{lstlisting}
def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): Ans @cps[Ans] = {
  val config = Config(e, env, store, time)
  if (cache.outContains(config)) Ans(cache.outGet(config), cache)
  else {
    val new_time = config.tick
    val new_cache = cache.outUpdate(config, cache.inGet(config))
    e match {
      case Let(x, App(f, ae), e) =>
        val closures = atomicEval(f, env, store)
        val (Clos(Lam(v, body), c_env), clscache) = choices[Storable](closures, new_cache)
        val vbaddr = allocBind(v, new_time)
        val new_cenv = c_env + (v -> vbaddr)
        val new_store = store.update(vbaddr, atomicEval(ae, env, store))
        /* evaluates the function application App(f, ae) */
        val Ans(bdvss, bdcache) = aeval(body, new_cenv, new_store, new_time, clscache)
        val (VS(vals, time, vsstore), vscache) = choices[VS](bdvss, bdcache)
        val baddr = allocBind(x, time)
        val new_env = env + (x -> baddr)
        val new_vsstore = vsstore.update(baddr, vals)
        /* evaluates the body expression from the let */
        val Ans(evss, ecache) = aeval(e, new_env, new_vsstore, time, vscache)
        Ans(evss, ecache.outUpdate(config, evss))

      case ae if isAtomic(ae) =>
        val vs = Set(VS(atomicEval(ae, env, store), new_time, store))
        Ans(vs, new_cache.outUpdate(config, vs))
    }
  }
}
\end{lstlisting}

@@

\begin{lstlisting}
def analyze(e: Expr) = {
  def iter(cache: Cache): Ans = {
    val Ans(vss, anscache) = reset { aeval(e, mtEnv, mtStore, mtTime, cache) }
    val initConfig = Config(e, mtEnv, mtStore, mtTime) 
    val new_cache = anscache.outUpdate(initConfig, vss)
    if (new_cache.out == cache.out) { Ans(vss, new_cache) }
    else { iter(Cache(new_cache.out, new_cache.out)) }
  }
  iter(mtCache)
}
\end{lstlisting}

>>>>

\begin{lstlisting}
val mtTime = List();                         val mtStore = Store[BAddr, Storable](Map())
val mtEnv = Map[String, BAddr]();            val mtKStore = Store[KAddr, Cont](Map(Halt -> Set()))
def inject(e: Expr): State = State(e, mtEnv, mtStore, mtKStore, Halt, mtTime)
\end{lstlisting}

@@

An instance of @VS@ represents the computational result of one path in
the nondeterministic evaluation.
The reason that we include a store is that there might be a case in which two paths
have the same values but different accumulated side effects (e.g., memory allocations).
Since the whole computation is potentially nondeterministic, the type of the final result
values is a collection @Set[VS]@. For convenience, we define it as a type @Ans@:

@@

An additional argument @continue@ which has type @Cont@ is also introduced.
Recall that in the disentangled AAM, the second case in @drive_step@
makes a call to @continue@. Here, the reemerging @continue@ is a function
of type @Ans => Ans@. The function @aeval@ does not return: instead, it calls
@continue@ with the result values.

@@

\begin{lstlisting}
def nd[T](ts: Iterable[T], acc: Ans, k: (T, Ans => Ans) => Ans, m: Ans => Ans): Ans = {
  if (ts.isEmpty) m(acc)
  else k(ts.head, (ans: Ans) => nd(ts.tail, acc ++ ans, k, m))
}
\end{lstlisting}

%TODO: better word to replace `accumulate`
Given a collection of elements of type @T@, an initial accumlated value of type 
@Ans@, a function @k@ that works on element's type @T@ and takes a continuation 
of type @Ans=>Ans@, and a function @m@ that works on and returns a value of type 
@Ans@ (which is also type @Ans=>Ans@), @nd@ chooses an element in the collection 
and applies @k@ to it, accumulates the value (@acc++ans@ where the @ans@ is the 
result comes from @k@) and eventually applies function @m@ on the accumulated 
value when there is no element in the collection.

%TODO: consider rewrite this paragraph and explain the relation to `continue`.
If we look at @continue@ and @ndcontinue@ in the disentangled AAM, we will find
that @nd@ is an abstraction of those two functions. In function @continue@,
there are two out calls to @drive_step@ and @ndcontinue@, here the application
of @k@ simulates @drive_step@, and @m@ simulates @ndcontinue@.

Readers familiar with functional programming may think of @fold@ at this point; indeed,
it is a variant implementation of @fold@ in continuation style. But here we have two
continuations, one for the next element in the collection, and one for the eventual result.
This is sometimes referred to as an extended continuation-passing style (ECPS)
that function @k@ is a normal continuation and @m@ is a meta-continuation 
\cite{Danvy:1990:AC:91556.91622}.

\textbf{Evaluation without Caching.}
Since our first version of @aeval@ implements only the nondeterministic evaluation, 
given the defined @nd@ operator, the evaluation function @aeval@ can be easily defined. 
In the case that @e@ is a @let@ expression, we invoke @nd@ twice and perform a nesting 
two-step, depth-first evaluation over possible values of @App(f, ae)@.
The values of branches from a fork point will be accumulated inside of @nd@, and 
then returned to the upper fork point.

\begin{lstlisting}
case Let(x, App(f, ae), e) =>
  val closures = atomicEval(f, env, store)
  nd[Storable](closures, Set[VS](), { case (clos, clsnd) =>
    val Clos(Lam(v, body), c_env) = clos
    val baddr = allocBind(v, new_time)
    val new_cenv = c_env + (v -> baddr)
    val new_store = store.update(baddr, atomicEval(ae, env, store))
    aeval(body, new_cenv, new_store, new_time, (bdvss: Set[VS]) => {
      nd[VS](bdvss, Set[VS](), { case (vs, bdnd) =>
        val VS(vals, time, store) = vs
        val baddr = allocBind(x, time)
        val new_env = env + (x -> baddr)
        val new_store = store.update(baddr, vals)
        /* evaluate e with new env and store, then transfer to next body's value */
        aeval(e, new_env, new_store, time, bdnd)
      }, clsnd) /* transfer to next target closure of f */
    })
  },
  continue)
\end{lstlisting}

The outer application of @nd@ evaluates over the set of possible target closures of
@f@. For each such closure, we go into its body with a new environment and a new store.
Remember that the function @aeval@ returns a set of @VS@ objects to its continuation,
so the continuation argument @bdvss@ is a set of @VS@ objects which represent values
from multiple computation paths when evaluating this single @body@ expression.

We next consider the inner application of @nd@, which evaluates over the set of body's 
value/stores @bdvss@. For each @VS@, the timestamp and store in @VS@ will be instantiated to
the next application of @aeval@. The inner application of @aeval@ evaluates @e@ with the
substituted environment and store, then returns a set of values of one computation path 
to its continuation @bdnd@, where the continuation will accumulate the result and 
transfer evaluation to the next body value.

The meta-continuation of the inner @nd@ is @clsnd@, meaning that we need to transfer 
to the next target closure of @f@. The meta-continuation of the outer @nd@ is just 
@continue@, as what we normally would do for CPS interpreters.

@@

If the argument expression @e@ is atomic, which is the second case of match, we simply 
construct a new instance of @VS@ and return it to the @continue@ function.

At the end, the initial invocation to @aeval@ is passed with initial environments,
especially an identity function as the halting continuation:
\begin{lstlisting}
def analyze(e: Expr) = aeval(e, mtEnv, mtStore, mtTime, (ans => ans))
\end{lstlisting}

@@

Based on the refunctionalized AAM from the last section, we use the same nondeterministic
operator @nd@, but further extend it to a fixed-point cached version in this section.
The fixed-point caching algorithm performs a sound over-approximation of all possible
concrete reachable paths and values, and also prevents non-termination when analyzing 
programs that may diverge.

@@

\textbf{Cache.}
Recall that the latent assumption of the worklist algorithm in classical AAM is that 
if we have seen a state @s@, it means we also have seen all the successors of state @s@.
The caching algorithm we adopt here replays the same assumption but in a big-step 
manner: if we have seen the configuration @c@, then it means we also have seen the 
values that are evaluated from the configuration @c@.
Here we will apply the fixed-point caching algorithm as described from
\citeauthor{darais2017abstracting}'s ADI paper \cite{darais2017abstracting}.

@@

\textbf{Cache-Passing Style.}
We can now transform our interpreter into cache-passing style. We begin by changing the 
return type @Ans@ to a case class which contains a @VS@ set and a cache. @Ans@ also supports
its own accumulating operation @++@ which defines as union on @Set[VS]@ and join on caches.

\begin{lstlisting}
case class Ans(vss: Set[VS], cache: Cache) {
  def ++(ans: Ans): Ans = { Ans(vss ++ ans.vss, cache.join(ans.cache)) }
}
\end{lstlisting}

The nondeterministic operator @nd@ is also slightly rewrite to cache-passing style:
continuation @k@ takes an additional argument of type @Cache@; the current accumulated
cache @acc.cache@ is exposed to @k@ when applies.

\begin{lstlisting}
def nd[T](ts: Iterable[T], acc: Ans, k: (T, Cache, Ans => Ans) => Ans, m: Ans => Ans): Ans = {
  if (ts.isEmpty) m(acc)
  else k(ts.head, acc.cache, (ans: Ans) => nd(ts.tail, acc ++ ans, k, m))
}
\end{lstlisting}

The skeleton of function @aeval@ largely remains the same.

\begin{lstlisting}
def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache, continue: Cont): Ans = {
  val config = Config(e, env, store, time)
  if (cache.outContains(config)) cont(Ans(cache.outGet(config), cache))
  else {
    val new_cache = cache.outUpdate(config, cache.inGet(config))
    val new_time = config.tick 
    e match { ...  }
  }
}
\end{lstlisting}

@@

The two cases of match are shown below. Note that the cache is used in a monotonic way: 
each function call to @aeval@ or @nd@ is passed with the latest cache from the most recent 
continuation of @aeval@ or @nd@.

\begin{lstlisting}
case Let(x, App(f, ae), e) =>
  val closures = atomicEval(f, env, store)
  nd[Storable](closures, Ans(Set[VS](), new_cache), {
    case (Clos(Lam(v, body), c_env), clscache, clsnd) =>
      val vbaddr = allocBind(v, new_time); val new_cenv = c_env + (v -> vbaddr)
      val new_store = store.update(vbaddr, atomicEval(ae, env, store))
      aeval(body, new_cenv, new_store, new_time, clscache, { case Ans(bdvss, bdcache) =>
        nd[VS](bdvss, Ans(Set[VS](), bdcache), {
          case (VS(vals, time, vssstore), bdvsscache, bdnd) =>
            val baddr = allocBind(x, time); val new_env = env + (x -> baddr)
            val new_store = vssstore.update(baddr, vals)
            aeval(e, new_env, new_store, time, bdvsscache, {
              case Ans(evss, ecache) => bdnd(Ans(evss, ecache.outUpdate(config, evss)))
            })
        }, clsnd)
      })
  },
  continue)
case ae if isAtomic(ae) =>
  val vs = Set(VS(atomicEval(ae, env, store), new_time, store))
  continue(Ans(vs, new_cache.outUpdate(config, vs)))
\end{lstlisting}

@@

\begin{lstlisting}
val mtCache = Cache(Store[Config, VS](Map()), Store[Config, VS](Map()))
def analyze(e: Expr) = {
  def iter(cache: Cache): Ans = {
    val Ans(vss, new_cache) = aeval(e, mtEnv, mtStore, mtTime, cache, { case Ans(vss, cache) => 
        Ans(vss, cache.outUpdate(Config(e, mtEnv, mtStore, mtTime), vss))
    })
    if (new_cache.out == cache.out) { Ans(vss, new_cache) }
    else { iter(Cache(new_cache.out, new_cache.out)) }
  }
  iter(mtCache)
}
\end{lstlisting}

@@

\textbf{Nondeterminism Operators, Again.}
%Introducing @Cache@ explicitly as the part of arguments of continuation @k@ reveals 
%that cache should be monotonically accumulated to the rest of computation. 
The first operator @nd@ is adapted from the previous one, but the meta-continuation
@m@ has been eliminated here. Now the only continuation @k@ applies on tuple type @(T, Cache)@ 
where the value of type @T@ comes from the collection, and @nd@ performs recursive invocation
on itself over elements in the collection @ts@.

@@

The function @choices@ takes two arguments: a set of elements with type @T@, 
and an initial accumulated value of type @Ans@.
The @choices@ iteratively returns an element from the collection along with the 
latest cache, and the return type is also annotated by \verb|@cps[Ans]|. 
This annotation denotes the final returned value of @choices@ is type @Ans@.
The @f@ is the delimited continuation captured by the @shift@ operator, which
is form by the statements after @choices@'s call-site.

Now we can use these two operators to redefine @aeval@.
We add an annotation \verb|@cps[Ans]| on the return type of the @aeval@ function to enable
applying delimited control operators.
Then, for every nondeterministic choice in the abstract interpreter, we can simply call 
@choices@ on the set of possible choices and write the program sequentially as concrete 
interpreters, and the subsequent statements and expressions become the delimited continuation
@f@. @choices@ also returns the latest cache to its left-hand side definition.
The cache is still accumulated to the next call to @choices@ or @aeval@, 
given the fact that the abstract interpreter should always use the latest cache.

\begin{lstlisting}
def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): Ans @cps[Ans] = {
  val config = Config(e, env, store, time)
  if (cache.outContains(config)) Ans(cache.outGet(config), cache)
  else {
    val new_time = config.tick; val new_cache = cache.outUpdate(config, cache.inGet(config))
    e match {
      case Let(x, App(f, ae), e) =>
        val closures = atomicEval(f, env, store)
        val (Clos(Lam(v, body), c_env), clscache) = choices[Storable](closures, new_cache)
        val vbaddr = allocBind(v, new_time); val new_cenv = c_env + (v -> vbaddr)
        val new_store = store.update(vbaddr, atomicEval(ae, env, store))
        val Ans(bdvss, bdcache) = aeval(body, new_cenv, new_store, new_time, clscache)
        val (VS(vals, time, vsstore), vscache) = choices[VS](bdvss, bdcache)
        val baddr = allocBind(x, time); val new_env = env + (x -> baddr)
        val new_vsstore = vsstore.update(baddr, vals)
        val Ans(evss, ecache) = aeval(e, new_env, new_vsstore, time, vscache)
        Ans(evss, ecache.outUpdate(config, evss))
      case ae if isAtomic(ae) =>
        val vss = Set(VS(atomicEval(ae, env, store), new_time, store))
        Ans(vss, new_cache.outUpdate(config, vss))
    }
  }
}
\end{lstlisting}

@@

\begin{lstlisting}
def analyze(e: Expr) = {
  def iter(cache: Cache): Ans = {
    val Ans(vss, anscache) = reset { aeval(e, mtEnv, mtStore, mtTime, cache) }
    val new_cache = anscache.outUpdate(Config(e, mtEnv, mtStore, mtTime), vss)
    if (new_cache.out == cache.out) { Ans(vss, new_cache) }
    else { iter(Cache(new_cache.out, new_cache.out)) }
  }
  iter(mtCache)
}
\end{lstlisting}
